Displacement of fishing effort by Large Scale Marine Protected Areas\footnote{Work in progress, do not circulate}

### 2018-10-21 docs/Manuscript.tex
Updated all

  \title{Displacement of fishing effort by Large Scale Marine Protected
Areas\footnote{Work in progress, do not circulate}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-19}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

@@

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, and include increases in
abundance, biomass, and diversity within the bounded regions. These
no-take zones may produce spillover effects, which provide fish for
outside areas. However, the economic and ecological implications of
displacing fishing effort are not yet fully understood. Novel data
products that track fishing effort at the vessel-level allow us to
identify changes in fleet- and vessel-level behavior upon the
implementation of protected areas, as well as how these redistribute.
This papers evaluates the implications of implementing LSMPA, by
evaluating changes in fishing hours, showing that vessels in the
effected region reduce fishing effort after the implementation of PIPA.
Our results are robust to a set of specifications. We also track the
relative spatial allocation of fishing events thorugh time, and identify
that areas closer to PIPA show an increase in relative fishing hourse
due to the displacement of PIPA-fishing vessels. Our results not only
provide an impact evaluation of the effect of LSMPAs on fishing
activity, but provide insights into vessel redistribution dynamics,
which may have ecological and economic implications.
\end{abstract}

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the later, and that are fully no take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}. The effectiveness and difficulties of
implementating LSMPAs are not the focus of this paper
\footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}.
Instead, we focus on the effect and implications of the ones that
\emph{have been} implemented.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protectes two
thrids of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of potential illegal shark fishing \emph{inside} a 2
million km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

@@

The empirical research that has been done in custumary sized MPAs
suggest that resource users may show idiosyncratic responses. For
example, \citet{stevenson_2013} show that a network of MPAs displaced
fishing effort farther away from ports, resulting in higher
\emph{perceived} costs, and increases in catch per unit effort.
\citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of a network of MPAs in
California, and find that commercial dive boats follow a
fishing-the-line pattern, while some fishing boats follow an ideal free
distribution. More recently \citet{elahi_2018} used satellite tracking
data to show that a temporal spatial closure caused trawlers to maintain
effort but apply it more intensively elsewhere, particularly along the
borders and closer to shore. The way in which fishers react to a spatial
closure can have major implications in its outcome
\citep{smith_2003,hilborn_2006} highlighting the need to understand how
fishers react to the implementation of LSMPAs, and fishing effort
changes and is spatially redistributed. However, all these closures took
place within Exlcusive Economic Zones, where other regulations exist.
This may not always be the case for LSMPAs, where often the entire EEZ
is converted into a LSMPA, leaving fishers with the options of moving to
the high seas or other countries' EEZs.

@@

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisors recomend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. A solid line indicates
countries that belong to the PNA, while a dashed line indicates all
others. A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. Theis region of the
Pacific has historically accounted for a large portion of tuna catches
\citep{aqorau_1997}. Today, the PNA controls close to 50\% of the global
skipjack tuna production \citep{pna_website_2018}. A large portion of
these catches derive from purse seine vessels licensed under the VDS.
Fishing vessels from Australia, New Zeland, China, France, Korea, Japan,
the Philippines, Taiwan, and the United States participate in the
purse-seining VDS.

@@

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore exlcuded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the anouncement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but are
only able to track 61 for the complete periof of study before and after
its
implementation\footnote{1 vessel never fished again after August 18, 2013.}.
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that
fished within PIPA at least once before the anouncement, and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure (n = 26).

We include three additional definitions of control groups as a
robustness check: one with only vessels that belong to PNA countries (n
= 7), and one that excludes Chinese vessels (n = 21). Our third control
is made up of Japanese purse seiners that fish in the Pacific but that
never fished inside PIPA (n =
27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify propper Taiwanese vessels, as I am currently using the entire Pacific ocean}.
Our definition of treatment and control groups leaves us with 61 and 21
treated and control vessels, which have just over 22 million
observations where about 22\% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain pannel
data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number
of vessels following a BACI design, as well as the fishing hours, before
and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours
for purse seiners have an abrupt increase, just before January 1st,
2015. This trend is observed for both treated and control groups. Across
all measures, the treatment and control vessels follow similar patterns,
confirming our claim that the control group provides a plausible
counterfactual. Figure \ref{fig:baci_strict} provides a visual
representation of the vessel-level fishing events that make up each
group through time. We use this data to answer our key questions: How do
the 61 purse seiners modify their behavior as compared to the different
control groups, and where do they go after the spatial closure? The
following section describes our empirical identification strategies.

@@

\caption{\label{tab:unnamed-chunk-6}\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Group & n & Before & After & Change (A / B)\\
\midrule
Control & 26 & 11.74 & 20.13 & 1.72\\
Treatment & 61 & 10.67 & 18.36 & 1.72\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}\label{fig:all_vessels}Fishing hours and
number of vessels by month for all vessels. Vertical dashed lines
indicates dates when satellites were added, solid line indicates PIPA
closurePIPA closure.}
\end{figure}

@@

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. A spatial closure might cause fishers to modify
their behavior as they adapt to a new state of the world. Some may have
to travel further distances to find new fishing grounds, which would
lead them to fish harder to compensate for the associated extra fuel
costs. If fishers had developed experience for fishing in particular
sites, being excluded might impose a learning cost on them, as they
identify new fishing grounds. This would result in increased search
times. However, if the area of influence of a vessel is significantly
larger than that of the spatial closure, they might already know other
places and times where to fish. Our variables of interests intend to
caputre these possible reactions.

@@

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month-, and flag-level dummies that account for seasonality or
country-level management
interventions\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

Our second part of the analyses focuses on the redistribution of fishing
effort. The role of institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season. If a vessel decides to fish within the PNA,
they've already made the decission to pay for access, expecting higher
returns from fishing there than in the High Seas. In the particular case
of PIPA, one would expect that a vessel holding a permit to fish in
I-Kiribati waters would then move to other fishing grounds within
Kiribati after the closure. Alternatively, if a vessel fishes illegaly
in I-Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other I-Kiribati waters, waters
of countries that have similar enforcement levels, or the High Seas.

We discretize spatial units by using a polygon for PIPA
\footnote{we would expect to see a decrease here} and distinct spatial
units for each EEZ of each country. Some vessels might shift from EEZs
into the high seas, but we are interested in knowing \emph{where} in the
high seas, so we incorporate additional regions by using a 1 degree
buffer of the high seas arround each of the EEZ regions. The rest of the
high seas are merged into a single spatial unit. For example, if we were
to do this only for Kiribati, we would have 8 spatial units: PIPA, three
EEZs, three 1-degree buffers of high seas arround each EEZs, and the
rest of the high seas. Whenever the buffers overlapped between
themselves, we randocmply clipped one on to the other. EEZs that had
sporadic fishing events were pooled into a group of ``others''.

To evaluate this change in effort allocation, we regress our variable of
interest (\emph{i.e.} fishing hours) on the interaction between a dummy
variable indicating the policy intervention and a dummy variable for
countries. This gives us the by-country change in proportional
allocation of fishing effort:

\[
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries for the spatial units defined above. Our parameter of
interest is \(\beta_{1,i}\), which captures the country-level change in
proportional fishing effort.

@@

\caption{\label{tab:unnamed-chunk-9}\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.}
\centering
\begin{tabular}[t]{lr}
\toprule
country & change\\
\midrule
EEZ COK 1 & 0.53\\
EEZ FSM 1 & 0.81\\
EEZ KIR 1 & -0.16\\
EEZ KIR 2 & 4.50\\
EEZ KIR 3 & -2.79\\
\addlinespace
EEZ MHL 1 & -0.44\\
EEZ NRU 1 & 0.07\\
EEZ PNG 2 & -9.07\\
EEZ SLB 1 & 3.18\\
EEZ TUV 1 & 1.27\\
\addlinespace
HS & 3.93\\
HS COK 1 & 0.06\\
HS KIR 1 & 2.96\\
HS KIR 2 & 1.07\\
HS KIR 3 & 2.53\\
PIPA PIPA 1 & -8.25\\
\bottomrule
\end{tabular}
\end{table}

@@

Our DiD analysis shows an overall increase in purse seine fishing hours,
even after accounting for the introduction of new satellites (Table
\ref{tab:did}). This coefficient estimate is consistent for different
model specifications and across groups of treatment and controls. These
effectively represent the patterns observed in Figure
\ref{fig:all_vessels}. The \(\beta_3\) coefficient indicating our
treatment effect suggests that, relative to the control, treated vessels
fish less, in the order of 0.5 hours per day. Another way to interpret
this is that the increase in fishing effort by treated vessels has
ocurred at a lower rate than control vessels. This result is robust and
significant (\(p < 0.05\)) for all model specifications using our main
control group, with the effect varying between \(\beta_3 = -0.515\) and
\(\beta_3 = -0.934\). These values are equivalent to a 15 - 28 reduction
in fishing hours per month.

Regressions coefficients for our DiD analysis are shown in Table
\ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use
only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels.
Columns 10 and 11 use Japanese vessels in the Pacific as controls, and
hence don't include flag FEs to avoid perfect colinearity between
Japanese flag and
control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}.
It must be noted, however, that when reducing the linear structure of
the \(Pre \times Post\) design and instead interacting the treatment
dummy with quarterly or year-month combinations we are not able to
reject the null hypotheses of no change (Figures
\ref{fig:q1,fig:q2,fig:q3,fig:q4,fig:ym1,fig:ym2,fig:ym3,fig:ym4}).

>>>>

  \title{Displacement of fishing effort by Large Scale Marine Protected
Areas\footnote{Work in progress, do not circulate}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-20}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

@@

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, but little empirical works
evaluate what happens to the effort after an MPA is created. The
economic and ecological implications of displacing fishing effort are
not yet fully understood. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analyses of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) Vessels that fished inside the protected area
redistribute to adjacent waters; and 3) We observe a crowding effect for
the first years after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications. As countries continue to implement
LSMPAs as a way to reach the stated 10\% target of ocean protection,
managers should consider how fishing effort will change in space and
through time to ensure that fishing effort is not just displaced
elsewhere, leading to overfishing in adjacent waters. While LSMPAs can
provide a wide range of benefits, their implementation must be
accompained with traditional fisheries management to maximize their
effectiveness.
\end{abstract}

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the later, and that are fully no take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}. The effectiveness and difficulties of
implementing LSMPAs are not the focus of this paper
\footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}.
Instead, we focus on the effect and implications of the ones that
\emph{have been} implemented.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{\label{fig:unnamed-chunk-13}\label{fig:LSMPAs_map}Large Scale
Marine Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protects two
thirds of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of potential illegal shark fishing \emph{inside} a 2
million km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

@@

The empirical research that has been done in customary sized MPAs
suggest that resource users may show idiosyncratic responses. For
example, \citet{stevenson_2013} show that a network of MPAs displaced
fishing effort farther away from ports, resulting in higher
\emph{perceived} costs, and increases in catch per unit effort.
\citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of a network of MPAs in
California, and find that commercial dive boats follow a
fishing-the-line pattern, while some fishing boats follow an ideal free
distribution. More recently \citet{elahi_2018} used satellite tracking
data to show that a temporal spatial closure caused trawlers to maintain
effort but apply it more intensively elsewhere, particularly along the
borders and closer to shore. The way in which fishers react to a spatial
closure can have major implications in its outcome
\citep{smith_2003,hilborn_2006} highlighting the need to understand how
fishers react to the implementation of LSMPAs, and fishing effort
changes and is spatially redistributed. However, all these closures took
place within Exclusive Economic Zones, where other regulations exist.
This may not always be the case for LSMPAs, where often the entire EEZ
is converted into a LSMPA, leaving fishers with the options of moving to
the high seas or other countries' EEZs.

@@

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisers recommend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{\label{fig:unnamed-chunk-14}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. This region of the
Pacific has historically accounted for a large portion of tuna catches
\citep{aqorau_1997}. Today, the PNA controls close to 50\% of the global
skipjack tuna production \citep{pna_website_2018}. A large portion of
these catches derive from purse seine vessels licensed under the VDS.
Fishing vessels from Australia, New Zealand, China, France, Korea,
Japan, the Philippines, Taiwan, and the United States participate in the
purse-seining VDS.

@@

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the announcement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but are
only able to track 61 for the complete period of study before and after
its
implementation\footnote{1 vessel never fished again after August 18, 2013.}.
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that
fished within PIPA at least once before the announcement, and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure (n = 26).

We include three additional definitions of control groups as a
robustness check: one with only vessels that belong to PNA countries (n
= 7), and one that excludes Chinese vessels (n = 21). Our third control
is made up of Japanese purse seiners that fish in the Pacific but that
never fished inside PIPA (n =
27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify proper Taiwanese vessels, as I am currently using the entire Pacific ocean}.
Our definition of treatment and control groups leaves us with 61 and 21
treated and control vessels, which have just over 22 million
observations where about 22\% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain panel
data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number
of vessels following a BACI design, as well as the fishing hours, before
and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours
for purse seiners have an abrupt increase, just before January 1st,
2015. This trend is observed for both treated and control groups. Across
all measures, the treatment and control vessels follow similar patterns,
confirming our claim that the control group provides a plausible
counterfactual. Figure \ref{fig:baci_strict} provides a visual
representation of the vessel-level fishing events that make up each
group through time. We use this data to answer our key questions: How do
the 61 purse seiners modify their behavior as compared to the different
control groups, and where do they go after the spatial closure? The
following section describes our empirical identification strategies.

@@

\caption{\label{tab:unnamed-chunk-17}\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Group & n & Before & After & Change (A / B)\\
\midrule
Control & 26 & 11.74 & 20.13 & 1.72\\
Treatment & 61 & 10.67 & 18.36 & 1.72\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
\caption{\label{fig:unnamed-chunk-18}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed lines
indicates dates when satellites were added, solid line indicates PIPA
closurePIPA closure.}
\end{figure}

@@

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. A spatial closure might cause fishers to modify
their behavior as they adapt to a new state of the world. Some may have
to travel further distances to find new fishing grounds, which would
lead them to fish harder to compensate for the associated extra fuel
costs. If fishers had developed experience for fishing in particular
sites, being excluded might impose a learning cost on them, as they
identify new fishing grounds. This would result in increased search
times. However, if the area of influence of a vessel is significantly
larger than that of the spatial closure, they might already know other
places and times where to fish. Our variables of interests intend to
capture these possible reactions.

@@

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \theta_t + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month- and flag-level dummies that account for seasonality or
country-level management interventions. \(\theta_t\) are dummies to
account for the addition of more satellites.
\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

Our second part of the analyses focuses on the redistribution of fishing
effort. The role of institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season. If a vessel decides to fish within the PNA,
they've already made the decision to pay for access, expecting higher
returns from fishing there than in the High Seas. In the particular case
of PIPA, one would expect that a vessel holding a permit to fish in
I-Kiribati waters would then move to other fishing grounds within
Kiribati after the closure. Alternatively, if a vessel fishes illegally
in I-Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other I-Kiribati waters, waters
of countries that have similar enforcement levels, or the High Seas.

We discretize spatial units by using a polygon for
PIPA\footnote{we would expect to see a decrease here} and distinct
spatial units for each EEZ of each country. Some vessels might shift
from EEZs into the high seas, but we are interested in knowing
\emph{where} in the high seas, so we incorporate additional regions by
using a 1 degree buffer of the high seas around each of the EEZ regions.
The rest of the high seas are merged into a single spatial unit. For
example, if we were to do this only for Kiribati, we would have 8
spatial units: PIPA, three EEZs, three 1-degree buffers of high seas
around each EEZs, and the rest of the high seas. Whenever the buffers
overlapped between themselves, we randomly clipped one on to the other.

We then take this spatial polygons and rasterize them to a 1-degree grid
(See Figure \ref{fig:raster_rgn}), which we then use to rasterize points
of fishing activity. For each cell, we calculate the total monthly
fishing hours by treated and control vessels (See Figure
\ref{fig:fishing_raster}). Since we are interested in the redistribution
of fishing effort, we use this gridded data to calculate the proportion
of fishing hours that each region receives each month relative to the
total fishing hours observed across all regions and all cells in that
month.

We use this measure to evaluate the change in effort allocation by
regressing it on the interaction between a year dummy and a dummy for
regions:

\[
y_{i,t} = \alpha + \beta_1Year_t + \beta_{2,i}Region_i + \beta_{3,i}Year_t \times Region_i+ \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that region \(i\) receives at time \(t\). Years are
modelled as factors (\(Year_t\)), using 2012 as the reference level.
\(Region\) is a dummy variable for regions defined above. Our parameter
of interest is \(\beta_{3,i}\), which captures the yearly by-country
change in proportional allocation of fishing effort relative to 2012.

@@

@@

Regressions coefficients for our DiD analysis are shown in Table
\ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use
only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels.
Columns 10 and 11 use Japanese vessels in the Pacific as controls, and
hence don't include flag fixed-effects to avoid perfect colinearity
between Japanese flag and
control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}.
Our DiD analysis shows an overall increase in purse seine fishing hours,
even after accounting for the introduction of new satellites (Table
\ref{tab:did}). This coefficient estimate is consistent for different
model specifications and across groups of treatment and controls, and
effectively represents the patterns observed in Figure
\ref{fig:all_vessels}.

The \(\beta_3\) coefficient indicating our treatment effect suggests
that, relative to the control, treated vessels fish less, in the order
of 0.5 hours per day. Another way to interpret this is that the increase
in fishing effort by treated vessels has occurred at a lower rate than
control vessels. The effect varying between \(\beta_3 = -0.515\) and
\(\beta_3 = -0.934\). These values are equivalent to a 15 - 28 reduction
in fishing hours per month. However, this result is only robust and
significant (\(p < 0.01\)) in the simplest specification of the
Diff-in-Diff. When reducing the linear structure of the
\(Pre \times Post\) design and instead interacting the treatment dummy
with quarterly or year-month combinations we are not able to reject the
null hypotheses of no change, suggesting that the total fishing hours
remains similar for the displaced vessels (Figs.
\ref{fig:q1}-\ref{fig:ym4}). Since vessels don't seem to be fishing less
(or at least not much less), an obvious question arises: where are they
fishing now?
